---
title: "Workflow for Reproducible Legacy Datasets in iotables"
format: html
editor: visual
---

This document explains how to rebuild, check, and maintain the legacy example datasets shipped in `iotables`. These datasets (e.g. `germany_1995`, `croatia_2010`, `uk_2010`) are used for unit tests and examples. They must be **long-term reproducible**.

## 1. Identify the canonical source

Each dataset has three representations:

-   **Manual example** (Eurostat, UN): the authoritative published table.
-   **CSV transcription** in `data-raw/`: clean copy of the manual table.
-   **CRAN dataset** in `data/`: the R data.frame shipped with the package.

The CSV + a builder script in `data-raw/` must always recreate the CRAN dataset.

## 2. Record reference metadata once

Run this once with the CRAN dataset to capture its structure:

```{r reference, eval=FALSE}
data(germany_1995, package = "iotables")

metadata <- list(
  col_classes = sapply(germany_1995, class),
  prod_na_order = as.character(unique(germany_1995$prod_na)),
  iotables_col_levels = levels(germany_1995$iotables_col),
  prod_na_lab_lookup = setNames(
    germany_1995$prod_na_lab,
    paste(germany_1995$prod_na, germany_1995$iotables_col)
  ),
  iotables_row_lookup = setNames(
    germany_1995$iotablesrow,
    paste(germany_1995$prod_na, germany_1995$iotablescol)
  )
)

saveRDS(metadata, file.path("data-raw", "germany_1995_metadata.rds"))
```

Repeat for each dataset. Commit the \*\_metadata.rds alongside its CSV.

## 3. Write a builder

Each dataset has a small function in `data-raw/`:

```{r builder, eval=FALSE}
build_germany_1995 <- function(
    path = file.path("data-raw", "Beutel_15_4_esa2010.csv")) {
  germany <- read.csv(path, stringsAsFactors = FALSE) %>%
    mutate(total = agriculture_group + industry_group + construction +
      trade_group + business_services_group + other_services_group)

  germany_long <- germany %>%
    pivot_longer(
      cols = c(
        agriculture_group, industry_group, construction,
        trade_group, business_services_group, other_services_group,
        total,
        final_consumption_households, final_consumption_government,
        gross_capital_formation, inventory_change, exports,
        total_final_use
      ),
      names_to = "iotables_col",
      values_to = "values"
    )

  # add recoded induse and metadata here …
}
```

## 4. Apply reference metadata

After pivoting:

-   Reorder rows with `prod_na_order`.

-   Set factor levels with `iotables_col_levels`.

-   Copy label strings with `prod_na_lab_lookup` and `iotables_row_lookup`.

-   Coerce column classes to `col_classes`.

## 5. Compare before overwrite

Always compare against the CRAN dataset before saving:

```{r comparison, eval=FALSE}
old <- germany_1995
new <- build_germany_1995()

stopifnot(identical(unclass(old), unclass(new)))
```

Only if this passes:

```{r, eval=FALSE}
usethis::use_data(new, overwrite = TRUE)
```

## 6. Lock with tests

Each dataset gets a regression test in tests/:

```{r testing, eval=FALSE}
test_that("germany_1995 rebuilds identically", {
  new <- build_germany_1995()
  expect_identical(unclass(germany_1995), unclass(new))
})
```

## 7. Document provenance

In the roxygen docstring of each dataset:

-   Cite the manual source (Eurostat / UN, table number).
-   Note that the dataset is rebuilt from the CSV in data-raw/.
-   State that metadata (labels, factor levels) is applied to guarantee reproducibility.

## 8. Principles

Never use column ranges (e.g. a:b) in `pivot_longer()`. Always list the 13 columns explicitly to avoid silent mismatches.

-   Numbers come from CSV/manual.
-   Labels and factors come from saved metadata.
-   Tests guarantee stability.

## 9. Using the template functions

To simplify maintenance of all legacy datasets, two generic helpers are provided in `data-raw/`. The first, `extract_METADATA_TEMPLATE()`, records the column classes, factor levels, and label lookups from an existing dataset on CRAN, and saves them as an `*.rds` file in `data-raw/`. This needs to be run only once per dataset. The second, `build_TEMPLATE()`, reads the corresponding CSV file, pivots it into tidy long form, applies the saved metadata, and outputs a data frame that should be byte-for-byte identical to the CRAN dataset. For each legacy dataset, copy these templates, replace `TEMPLATE` with the dataset name (e.g. `germany_1995` or `croatia_2010_1700`), and adjust the CSV path, column list to pivot, and any dataset-specific metadata such as `geo` or `time`. This workflow guarantees reproducibility and avoids the trial-and-error we encountered when rebuilding Germany 1995.
