---
title: "terminology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{terminology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(scipen=999)
```



```{r setup}
library(iotables)
```

```{r}
germany_siot <- iotable_get()
```

### First quadrant

The first quadrant of the symmetric input-output tables, or in other words, the _domestic intermediate matrix_, input-output flows between a simplifed, 6 sectors of the German economy in 1995. 

```{r}
germany_siot[c(1:7), c(1:7)]
```

Rowwise we see the flow of inputs. For example, at basic prices, out of the total demand for business services(423933), 193176	of the production of the business services are used by other business service providers, 65755	by trade enterprises, and 34223 by other service industries.

```{r outputflow}
germany_siot[5,c(1, 5:8)]
```

The columns describe the structure of the input of the corresponding sector. For example, the business service sectors needed 15058 input from the other services, and 13371 imports from abroad (shown in the table below), to meet the final domestic demand of 423933 for business services (shown in the table above):


```{r inputflow}
germany_siot[c(5:9), c(1, 6)]
```

## Coefficients

Input coefficients for domestic intermediates

The `input_coefficient_matrix_create()` function relies on the following equation. The numbering of the equations is the numbering of the Eurostat Manual.

(9)  $a_{ij}$ = $x_{ij}$ / $x_j$  [recap: (43) is the same]

$a_{ij}$: input coefficient for domestic goods and services</br>
$x_{ij}$: flow of domestic commodity i to sector j, i.e., a cell in the first quadrant.</br>
$x_j$: output of j

```{r}
input_coefficient_matrix_create_2(iotable_get(), digits = 2)
```


Output coefficients
The output coefficients are the quotients of the quadrant I (intermediates) and the quadrant II (final demand elements) of a sector.

The `output_coefficient_matrix_create()` function creates these coefficients based on equation (5) in the Eurostat Manual.

(5) $o_{ij}$ = $x_{ij}$ / $x_i$ 
 
$o_{ij}$ = output coefficient for domestic goods and services (i = 1, ..., 6; j = 1, ..., 6)
$x_{ij}$ = flow of commodity i to sector j
$x_j$ = output of sector i


## Interindustry analysis

The term ‘forward linkage’ is used to indicate this interconnection of a particular sector to those to which it sells its output. In its simplest form, the strength of the forward linkage of a sector i is given by the rowwise sum of the direct output coefficients, which considers the direct supply-side economic affects of hte sector i on the entire domestic economy.

In its simplest form, the strength of the _backward linkage_ of a sector j is given by the column sum of the direct input coefficients, which only considers the direct, demand-side affects of sector j on the entire domestic economy. The sum of the inverse of the input coefficients also considers the indirect effects.

